---
title: "Brooke Final Data Analysis, Modeling_You Been"
author: "You Been PArk"
date: "2024-11-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Visualization
We first took an attempt at the visualization to see the bigger picture of the 
dataset provided. 


```{r}
# Load necessary libraries
library(tidyverse)

# Load the dataset
data <- read.csv("Individual-level basic variable.csv")

# Ensure Age and Partners_count are numeric
data$Age <- as.numeric(data$Age)
data$Partners_count <- as.numeric(data$Partners_count)
```

To begin with, we looked into the mean number of partners by age. 
```{r}
# Load necessary libraries
library(tidyverse)

# Load the dataset
data <- read.csv("Individual-level basic variable.csv")

# Ensure Age and Partners_count are numeric
data$Age <- as.numeric(data$Age)
data$Partners_count <- as.numeric(data$Partners_count)

# Group by Age and calculate the mean of Partners_count
age_partner_mean <- data %>%
  group_by(Age) %>%
  summarise(mean_partners = mean(Partners_count, na.rm = TRUE))

# Create a bar plot
ggplot(age_partner_mean, aes(x = Age, y = mean_partners)) +
  geom_bar(stat = "identity") +
  labs(title = "Mean Partners Count by Age", 
       x = "Age", 
       y = "Mean Partners Count") +
  theme_minimal()
```
With the data set given, which was an integer, the bar graph shows an increasing number of mean partner counts. So, we wanted to investigate more on age as a variable that will impact the number of partner counts.

After this, we took the idea from our client Brooke, and took a look at how each gender shows difference in the number of playmates.

```{r}
data$Sex <- as.factor(data$Sex)  # Make sure Sex is a factor (M/F)

# Group by Age and Sex, then calculate the mean of Partners_count
age_sex_partner_mean <- data %>%
  group_by(Age, Sex) %>%
  summarise(mean_partners = mean(Partners_count, na.rm = TRUE), .groups = "drop")

# Ensure all combinations of Age and Sex are present, even if no data exists
age_sex_full <- age_sex_partner_mean %>%
  tidyr::complete(Age, Sex, fill = list(mean_partners = NA))

# Create a bar plot with divided bars for Males and Females
ggplot(age_sex_full, aes(x = Age, y = mean_partners, fill = Sex)) +
  geom_bar(stat = "identity", position = "dodge", na.rm = FALSE) +
  labs(title = "Mean Partners Count by Age and Sex", 
       x = "Age", 
       y = "Mean Partners Count", 
       fill = "Sex") +
  theme_minimal()


```
As shown in the graph, there are more playmates for male than female except for 9 years old. While we do not have the data for 6 year old female and 17 year old male, we can see that male does have more playmates than that of female. 

We then divide the ages into two groups of Pre-pubescent and Post-pubescent which are below and above 12. . 

```{r}

# Group by Puberty_Status and calculate the mean of Partners_count
puberty_partner_mean <- data %>%
  group_by(Puberty_Status) %>%
  summarise(mean_partners = mean(Partners_count, na.rm = TRUE))

# Create a bar plot with the specified order
ggplot(puberty_partner_mean, aes(x = Puberty_Status, y = mean_partners, fill = Puberty_Status)) +
  geom_bar(stat = "identity") +
  labs(title = "Mean Partners Count by Puberty Status", 
       x = "Puberty Status", 
       y = "Mean Partners Count") +
  theme_minimal() +
  scale_fill_manual(values = c("Pre-pubescent" = "lightblue", "Post-pubescent" = "lightgreen"))

```
Generally, there are a bigger mean partners count for post-pubescent than pre-pubescent meaning that there is an increase in networks when they are older than 12. 

```{r}
pre_pubescent_data <- data %>% filter(Puberty_Status == "Pre-pubescent")
post_pubescent_data <- data %>% filter(Puberty_Status == "Post-pubescent")

# Group by Age and calculate the mean of Partners_count for both groups
pre_pubescent_mean <- pre_pubescent_data %>%
  group_by(Age) %>%
  summarise(mean_partners = mean(Partners_count, na.rm = TRUE))

post_pubescent_mean <- post_pubescent_data %>%
  group_by(Age) %>%
  summarise(mean_partners = mean(Partners_count, na.rm = TRUE))

# Create the bar plots for both groups

# Pre-pubescent plot
pre_plot <- ggplot(pre_pubescent_mean, aes(x = Age, y = mean_partners)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Mean Partners Count by Age (Pre-pubescent)", 
       x = "Age", 
       y = "Mean Partners Count") +
  theme_minimal() +
  ylim(0, 6)

# Post-pubescent plot
post_plot <- ggplot(post_pubescent_mean, aes(x = Age, y = mean_partners)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "Mean Partners Count by Age (Post-pubescent)", 
       x = "Age", 
       y = "Mean Partners Count") +
  theme_minimal()

# Print the plots
print(pre_plot)
print(post_plot)


```

Now, let us look at the caregiver's residency duration and how that could impact the number of playmates. 
We decided to split the dataset into people who have not immigrated, people who have immigrated in the last 10 years, and people who have immigrated more than 10 years ago. Since the CaregiverResidencyDuration show the immigration duration in months, we split the data between below 120 and 120 or above. 

```{r}
# Ensure Partners_count and CaregiverResidencyDuration are numeric
df$Partners_count <- as.numeric(df$Partners_count)
df$CaregiverResidencyDuration <- as.numeric(df$CaregiverResidencyDuration)

# Create a new variable for CaregiverResidencyDuration categories
df <- df %>%
  mutate(ResidencyCategory = ifelse(CaregiverResidencyDuration < 120, 'Below 120 months', '120 months or Above'))

# Set the factor levels to ensure the desired order
df$ResidencyCategory <- factor(df$ResidencyCategory, levels = c("Below 120 months", "120 months or Above"))

# Group by ResidencyCategory and calculate the mean of Partners_count
residency_partner_mean <- df %>%
  group_by(ResidencyCategory) %>%
  summarise(mean_partners = mean(Partners_count, na.rm = TRUE))

# Create the bar plot with the specified order
ggplot(residency_partner_mean, aes(x = ResidencyCategory, y = mean_partners, fill = ResidencyCategory)) +
  geom_bar(stat = "identity") +
  labs(title = "Mean Partners Count by Caregiver Residency Duration",
       x = "Caregiver Residency Duration",
       y = "Mean Partners Count") +
  theme_minimal() +
  scale_fill_manual(values = c("Below 120 months" = "lightblue", "120 months or Above" = "lightgreen"))

```

The NA values mostly represent the people who have been in this island without immigrating from another. The graph shows that the mean partners count is greatest for the people who have not immigrated, and still greater for a family who have immigrated when their children were very young, or not born yet. 


#Summary
From the visualization, we have acknowledged the fact that age, gender, maturity, and immigration duration could have an impact on the partner's count. Hence, when building a model, we will be investigating these as essential variables for consideration.

###Poisson Regression
1. The question we will take a look at is "How does immigration impact the number of playmates?"
```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(MASS)
library(AER)

# Load the dataset
data <- fread("Individual-level basic variable.csv")

# Filter out the 16-year-old girl
data <- subset(data, !(Age == 17 & Sex == "F"))

# Remove rows with NA values in variables used in the model
data <- data %>%
  filter(
    !is.na(Partners_count) &
    !is.na(Age) &
    !is.na(Sex) &
    !is.na(CaregiverImmigrate) &
    !is.na(CaregiverResidencyDuration)
  )

# Convert categorical variables to factors
data[, Sex := as.factor(Sex)]
data[, CaregiverImmigrate := as.factor(CaregiverImmigrate)]
data[, UserLanguage := as.factor(UserLanguage)]


```
We started with one of the most simple models, only including age, gender and CargiverImmigrate.

```{r}
# Fit the Poisson regression model after cleaning
poisson_model <- glm(
  Partners_count ~ Age + Sex + CaregiverImmigrate,
  family = poisson(link = "log"),
  data = data
)

summary(poisson_model)

dispersiontest(poisson_model)

# Observed counts
observed_counts <- table(data$Partners_count)

# Predicted (Expected) counts from the model
predicted_counts <- predict(poisson_model, type = "response")

# Get unique counts from observed data
unique_counts <- 0:max(data$Partners_count)

# Compute expected counts for all unique counts
expected_curve <- data.frame(
  count = unique_counts,
  expected = dpois(unique_counts, lambda = mean(predict(poisson_model, type = "response")))
)

# Add square root of expected counts for plotting
expected_curve$sqrt_expected <- sqrt(expected_curve$expected * nrow(data))

library(ggplot2)

ggplot(rootogram_data, aes(x = count)) +
  geom_bar(aes(y = sqrt_observed), stat = "identity", fill = "blue", alpha = 0.5) +
  geom_line(data = expected_curve, aes(x = count, y = sqrt_expected), color = "red", size = 1.2) +
  geom_segment(
    aes(x = count, xend = count, y = sqrt_observed, yend = sqrt_expected),
    color = "black", linetype = "dashed"
  ) +
  theme_minimal() +
  labs(
    title = "Rootogram with Curved Expected Line",
    x = "Count",
    y = "Square Root of Frequency"
  )



```
In this model, there is a slight over dispersion, with the value being 1.125. In a Poisson regression, the variance of the response variable is assumed to be equal to the mean. But, when the variance is greater than the mean, we say that there is an over dispersion. Then, we are expected to do some other tests to validate the poisson regression. Here, we have created a rootogram, which shows the square root of frequency for each count. The red line represents the expected frequency which is based on the poisson regression we have done above. When we compare the line of the expected frequency and the actual count from the model(shown in a bar graph), we can see that it does kind of follow the trend, but has some outliers such as 1 count and 8 count. Hence, we would like to investigate if there are any better models to see the more accurate relevance. 


So, what I decided to do was to add the caregiver immigration duration as well. 

```{r}
poisson_model2 <- glm(
  Partners_count ~ Age + Sex + CaregiverImmigrate + CaregiverResidencyDuration ,
  family = poisson(link = "log"),
  data = data
)

summary(poisson_model2)

dispersiontest(poisson_model2)

# Fit the Poisson regression model
poisson_model2 <- glm(
  Partners_count ~ Age + Sex + CaregiverImmigrate + CaregiverResidencyDuration,
  family = poisson(link = "log"),
  data = data
)

# Model summary
summary(poisson_model2)

# Test for overdispersion
library(AER)
dispersiontest(poisson_model2)

# Fit the Poisson regression model
poisson_model2 <- glm(
  Partners_count ~ Age + Sex + CaregiverImmigrate + CaregiverResidencyDuration,
  family = poisson(link = "log"),
  data = data
)

# Model summary
summary(poisson_model2)

# Test for overdispersion
library(AER)
dispersiontest(poisson_model2)


```
By adding another variable, we can see that the dispersion has fallen to 1.10, but still has similar look when you consider the rootogram. The standard error, which represents how big of an error it contains considering the coefficient, is also still big, which leads to us to saying this has not improved the model as much. 

One of the other things that our client Brooke has mentioned was use of language. Language is also a big factor considering the immigration status. So, we will see if this also does impact the number of playmates. Also, I have added an interaction term of user language and caregiver residency duration in order to see if the effect of caregiver residency duration on partners count depends on the value of user language. 
```{r}
poisson_model3 <- glm(
  Partners_count ~ Age + Sex + CaregiverImmigrate + CaregiverResidencyDuration + UserLanguage +
    CaregiverResidencyDuration:UserLanguage,
  family = poisson(link = "log"),
  data = data
)

summary(poisson_model3)

dispersiontest(poisson_model3)

# Observed counts
observed_counts <- table(data$Partners_count)

# Predicted (Expected) counts from the model
predicted_counts <- predict(poisson_model3, type = "response")

# Get unique counts from observed data
unique_counts <- 0:max(data$Partners_count)

# Compute expected counts using individual predicted lambdas
expected_curve <- data.frame(
  count = unique_counts,
  expected = sapply(unique_counts, function(x) sum(dpois(x, lambda = predicted_counts)))
)

# Add square root of expected counts for plotting
expected_curve$sqrt_expected <- sqrt(expected_curve$expected)

# Create rootogram data with observed counts
rootogram_data <- data.frame(
  count = as.numeric(names(observed_counts)),
  observed = as.numeric(observed_counts),
  sqrt_observed = sqrt(as.numeric(observed_counts))
)

# Merge rootogram data with expected curve data
rootogram_data <- merge(
  rootogram_data,
  expected_curve,
  by = "count",
  all.x = TRUE
)

# Plot the rootogram
library(ggplot2)

ggplot(rootogram_data, aes(x = count)) +
  geom_bar(aes(y = sqrt_observed), stat = "identity", fill = "blue", alpha = 0.5) +
  geom_line(aes(y = sqrt_expected), color = "red", size = 1.2) +
  geom_segment(
    aes(x = count, xend = count, y = sqrt_observed, yend = sqrt_expected),
    color = "black", linetype = "dashed"
  ) +
  theme_minimal() +
  labs(
    title = "Rootogram with Curved Expected Line",
    x = "Count",
    y = "Square Root of Frequency"
  )

predicted_counts_model1 <- predict(poisson_model2, type = "response")
predicted_counts_model3 <- predict(poisson_model3, type = "response")

summary(predicted_counts_model1)
summary(predicted_counts_model3)


```

From the model summary, we see that there has been some significant improvements in the model by adding these variables. From the interaction term summary, we can see that the effect of caregiver residency duration on partners count is slightly positive for Spanish speakers. From the summary se wee that some of the variables here show significance, such as sex, user language, and the interaction term. The coeffecient estimate is significantly bigger than the standard error, which implies that the estimation made by the poisson model does not hold a big error. 
The over dispersion rate has fallen from 1.125 to 1.055 which is closer to 1, and the AIC which is better when smaller has also decresed from 420.42 to 419.37. We have added a summary of the two models, model 1 and model 3, showing how the predicted minimum and maximum. As you can see, the maximum has grown to 8.679 from 4.884 showing that now, the model is putting bigger partner counts into account. 


We will now run a quassi poisson model, to see if the over dispersion will be better solved, and improved the model. 
 
```{r}

quasipoisson_model3 <- glm(
  Partners_count ~ Age + Sex + CaregiverImmigrate + CaregiverResidencyDuration + UserLanguage +
    CaregiverResidencyDuration:UserLanguage,
  family = quasipoisson(link = "log"),
  data = data
)

# Summary of the quasi-Poisson model
summary(quasipoisson_model3)

# Calculate the dispersion parameter
dispersion <- sum(residuals(poisson_model3, type = "pearson")^2) / df.residual(poisson_model3)
dispersion


```

The over dispersion rate has increased to 1.135, showing that quassi poisson modeling worsens the dispersion. 

```{r}

library(MASS)
negbin_model3 <- glm.nb(
  Partners_count ~ Age + Sex + CaregiverImmigrate + CaregiverResidencyDuration + UserLanguage +
    CaregiverResidencyDuration:UserLanguage,
  data = data
)

# Summary of the Negative Binomial model
summary(negbin_model3)

# Dispersion parameter (theta)
negbin_model3$theta


```
Even from the negative binomial modeling with the same variables, we see an increase in standard error, compared to the standard error of poisson model 3. This shows that we can conclude that model 3 is better than the other two models, with the standard errors and over dispersion being smaller. 

#Summary of modeling
We have conducted several poisson models with the variables that were considerable as affecting the number of playmates. From the investigation starting with a simple model to a more complicated model, we could conclude that 

Partners_count ~ Age + Sex + CaregiverImmigrate + CaregiverResidencyDuration + UserLanguage + CaregiverResidencyDuration:UserLanguage

shows the best fit to examine the data. 

We can see that Sex, Language, and CaregiverResidancyDuration:Language is statistically significant, meaning these are some of the most important variables that impact the number of partners count. 

We can interpret the coefficients this way. Boys tend to have exp(0.183) increase in playmates than that of girls, with the error around 0.108 for the coefficients. Children who use Spanish tend to have exp(-1.191 ) decrease in playmates than who speak English with the error range of around 0.524. The interaction term shows that every month increase in the residency duration leads to an increase in exp(0.002568) of playmates, when the child uses Spanish. Hence, children who do speak Spanish will find more friends according to the increase in the months they stay in the Island. 


