---
title: "Louis_individual_analysis"
output: pdf_document
date: "2024-10-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load the necessary libraries
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
# Load the datasets
df_basic <- read.csv("Individual-level basic variable.csv")
df_calculate <- read.csv("Individual-level with calculate.csv")

# View the structure of the data
str(df_basic)
str(df_calculate)
```

## handle the missing data
```{r}
# Check for missing values
colSums(is.na(df_basic))
colSums(is.na(df_calculate))

df_basic_clean <- df_basic %>% drop_na()
df_calculate_clean <- df_calculate %>% drop_na()

summary(df_basic_clean)
summary(df_calculate_clean)
```

## Age and Sex Distribution
```{r}
ggplot(df_basic, aes(x = Age)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Age Distribution", x = "Age", y = "Frequency")
```

A little right skewed.

```{r}
ggplot(df_basic, aes(x = Sex)) +
  geom_bar(fill = "green") +
  theme_minimal() +
  labs(title = "Sex Distribution", x = "Sex", y = "Count")
```
Number of female and male playmates are close.

```{r}
# Simple linear regression using Age as the predictor for Count_unrelated
model <- lm(Count_unrelated ~ Age, data = df_basic)
summary(model)

ggplot(df_basic, aes(x = Age, y = Count_unrelated)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  theme_minimal() +
  labs(title = "Linear Regression: Age vs. Count of Unrelated Individuals", x = "Age", y = "Count Unrelated")
```

## check for other possible predictors
```{r}
# Select relevant numeric and factor variables for the analysis
df_basic_clean <- df_basic %>% select(Age, Sex, SchoolSimplified, Neighborhood, Count_unrelated)

# Check for correlations (only for numeric variables)
cor(df_basic_clean %>% select_if(is.numeric))

summary(df_basic_clean)
```

```{r}
# Convert categorical variables to factors
df_basic_clean$Sex <- as.factor(df_basic_clean$Sex)
df_basic_clean$SchoolSimplified <- as.factor(df_basic_clean$SchoolSimplified)
df_basic_clean$Neighborhood <- as.factor(df_basic_clean$Neighborhood)

# Fit a multiple linear regression model
model_multi <- lm(Count_unrelated ~ Age + Sex + SchoolSimplified + Neighborhood, data = df_basic_clean)

summary(model_multi)
```
1. Age is statistically significant.

2. Neighborhood: Children living in the Countryside are expected to know around 6.39 more unrelated individuals than children in the reference neighborhood, indicating that the Countryside neighborhood has a strong positive effect on the number of unrelated playmates.

## improve the model
```{r}
# Fit a refined model with only Age and Neighborhood as predictors
model_refined <- lm(Count_unrelated ~ Age + Neighborhood, data = df_basic_clean)

summary(model_refined)
```

```{r}
# Fit a model with interaction between Age and Neighborhood
model_interaction <- lm(Count_unrelated ~ Age * Neighborhood, data = df_basic_clean)

# Summarize the model with interaction
summary(model_interaction)
```

It failed to improve the situation.